\section{Two-phase consolidation}

For the two-phase system the governing equations presented in the previous sections must be expanded slightly.

\subsubsection*{Fluid mass and momentum balance}
Following the same mass balance procedure as for the single phase case, and gathering two mass balance equations, one for a wetting fluid (i.e. liquid) and one for a non-wetting (i.e. gas) fluid, we write,

\begin{equation}
\begin{split}
{{S}_{w}}\left( \frac{\phi }{{{K}_{w}}}+\frac{\alpha -\phi }{{{K}_{s}}} \right)\frac{d{{p}_{w}}}{dt}-\phi \frac{d{{S}_{nw}}}{dt}+... \\
\nabla \cdot \left\{ \frac{\mathbf{k}k_{w}^{r}}{{{\mu }_{w}}}\left( -\nabla {{p}_{w}}+{{\rho }_{w}}\mathbf{g} \right) \right\}+\alpha {{\text{S}}_{w}}\nabla \cdot \frac{\partial \mathbf{u}}{\partial t}=0
\end{split}
\label{eq:H2liq}
\end{equation}

for mass balance of the wetting fluid, subscript $w$, and,

\begin{equation}
\begin{split}
{{S}_{nw}}\left( \frac{\phi }{{{K}_{nw}}}+\frac{\alpha -\phi }{{{K}_{s}}} \right)\frac{d{{p}_{w}}}{dt}+\phi \left( 1-\frac{{{S}_{nw}}}{{{K}_{nw}}}\frac{\partial {{p}_{c}}}{\partial {{S}_{w}}} \right)\frac{d{{S}_{nw}}}{dt}+... \\
\nabla \cdot \left\{ \frac{\mathbf{k}k_{nw}^{r}}{{{\mu }_{nw}}}\left( -\nabla {{p}_{w}}-\frac{\partial {{p}_{c}}}{\partial {{S}_{w}}}\nabla {{S}_{nw}}+{{\rho }_{nw}}\mathbf{g} \right) \right\}+\alpha {{S}_{nw}}\nabla \cdot \frac{\partial \mathbf{u}}{\partial t}= 0
\end{split}
\label{eq:H2gas}
\end{equation}

for mass balance of the non-wetting fluid, subscript $nw$. In these equations, $k^{r}_{\alpha }$ is relative permeability of phase $\alpha $, $\rho _{\alpha }$ is density of phase $\alpha $, and we have chosen wetting pressure, $p_w$, and non-wetting saturation, $S_{nw}$, as primary variables in the solution scheme. Other primary variables, such as capillary pressure, $p_c$, and non-wetting pressure, $p_{nw}$, could also have been chosen with algebraic manipulation, but our benchmark example requires constant, $p_c=0$, capillary pressure for comparison with an anylitical solution. This is not possible, of course, if $p_c$ is a primary variable. Any viable permeability saturation function may be chosen for the example, we use the Brooks-Corey function.

For the numerical solution, storage due to two different fluids with two different compressibilities and densities is handled implicitly with solution of the above equations. For the analytical solution, we must define an effective compressibility as a function of fluid saturation and properties of each fluid. For immiscible fluids without penetrating bubbles, two compressible materials behave as resistors in series with respect to bulk modulus, thus the effective modulus is

\begin{equation}
\frac{1}{{{K}_{f}}}=\frac{{{S}_{w}}}{{{K}_{w}}}+\frac{{{S}_{nw}}}{{{K}_{nw}}}.
\end{equation}

\subsubsection*{Solid momentum balance}

As with the single phase case, balance of linear momentum is defined by,

\begin{equation}
\frac{\partial {{\sigma }_{ij}}}{\partial {{x}_{j}}}+{{F}_{i}}=0,
\end{equation}

but now the body force is a function of two fluids, $F=\rho _m \text{g}$ for the mixture density $\rho _m =\phi (S_w \rho _w + S_{nw} \rho _{nw}) + (1-\phi )\rho _s$, and insertion of the elastic consitutive law yields for solid displacement,

\begin{equation}
\frac{\partial }{\partial {{x}_{j}}}\left[ G\frac{\partial {{u}_{i}}}{\partial {{x}_{j}}}+\left( \lambda +G \right)\frac{\partial {{u}_{j}}}{\partial {{x}_{i}}}-\alpha \bar{p}{{\delta }_{ij}} \right]+{{F}_{i}}=0,
\end{equation}

where \textit{mean} fluid pressure is defined as $\bar{p}=S_wp_w + S_{nw}p_{nw}$.

\subsubsection*{FEM solution scheme}
For the two-phase simulations, the two fluid balance equations are solved in a single global equation, with an iterative coupling to the solid balance equation. 

%-------------------------------------------------------------------------
%-------------------------------------------------------------------------

\input{chapter_14/subsection_14_3_1}
\input{chapter_14/subsection_14_3_2}
\input{chapter_14/subsection_14_3_3}
