\section{Single phase consolidation}
Let us begin by introducing a few governing equations.

\subsubsection*{Fluid mass and momentum balance}
Linear momentum balance for the fluid follows Darcy:

\begin{equation}
{v}_{ri}=\phi \left( {{v}_{i}}-{{v}_{si}} \right)=-\frac{{{k}_{ij}}}{\mu }\left( \frac{\partial p}{\partial {{x}_{j}}}+\rho {{g}_{j}} \right),
\label{eqch14:genmass}
\end{equation}

for the intrinsic permeability, ${{k}_{ij}}$, dynamic viscosity, $\mu $, and density, $\rho $. The subscript $r$ considers that fluid velocity is relative to motion of the deformable solid (${{v}_{si}}$), so that ${{v}_{i}}$ is absolute fluid velocity, and ${{v}_{ri}}$ is relative. Conservation of fluid mass requires,

\begin{equation}
\frac{\partial }{\partial t}\left( \phi \rho  \right)+\frac{\partial }{\partial {{x}_{i}}}\left( \phi \rho {{v}_{i}} \right)=0.
\end{equation}
	
Fluid properties are functions of temperature and pressure. The fluid density time derivative appearing in the mass balance equation may be expanded to

\begin{equation}
\frac{d\rho}{dt}=\rho \left( \frac{1}{K_{f}^{p}}\frac{dp}{dt}-\frac{1}{K_{f}^{T}}\frac{dT}{dt} \right),
\end{equation}
	
with fluid compressibility given by $1/K_{f}^{p}={{\left. \left( 1/\rho  \right)\left( \partial \rho /\partial p \right) \right|}_{T}}$ and for the fluid thermal expansion coefficient $1/K_{f}^{T}=-{{\left. \left( 1/\rho  \right)\left( \partial \rho /\partial T \right) \right|}_{p}}$. In these definitions we utilize moduli ($K$ = inverse compressibility). Also, because thermal effects are not considered in these examples, the temperature dependence may be neglected. Utilizing the Lagrangian total derivative of a component relative to the moving solid, ${{d}_{s}}/dt=\partial /\partial t+{{v}_{si}}\partial /\partial {{x}_{i}}$, and a moving fluid, ${{d}_{f}}/dt=\partial /\partial t+{{v}_{fi}}\partial /\partial {{x}_{i}}$, substituting for absolute fluid velocity and dividing through by density gives,

\begin{equation}
\phi \left( \frac{1}{K_{f}^{p}}\frac{dp}{dt}+\frac{{{d}_{s}}\phi }{dt}+\frac{\partial {{v}_{si}}}{\partial {{x}_{i}}} \right)=-\frac{\partial }{\partial {{x}_{i}}}\left( {{v}_{ri}} \right).
\end{equation}
	
To obtain the porosity time derivative, we expand the solid mass balance to obtain

\begin{equation}
\frac{{{d}_{s}}\phi }{dt}=\frac{\left( 1-\phi  \right)}{{{\rho }_{s}}}\frac{\partial {{\rho }_{s}}}{\partial t}+\left( 1-\phi  \right)\frac{\partial {{v}_{si}}}{\partial {{x}_{i}}}.
\end{equation}
	
Substituting this gives,

\begin{equation}
\frac{\phi }{K_{f}^{p}}\frac{dp}{dt}+\left[ \frac{\partial {{v}_{si}}}{\partial {{x}_{i}}}+\frac{\left( 1-{{\rho }_{s}} \right)}{{{\rho }_{s}}}\frac{\partial {{\rho }_{s}}}{\partial t} \right]=-\frac{\partial }{\partial {{x}_{i}}}\left( {{v}_{ri}} \right).
\end{equation}
	
Utilizing Biot's formulation to represent the solid density time derivative and assuming small strain yields the full fluid mass balance (\cite{KhaliliSelvadurai:03} and \cite{RutEtAl:01}),

\begin{equation}
\left( \frac{\phi }{K_{f}^{p}}+\overbrace{\frac{\left( \alpha -\phi  \right)}{{{K}_{g}}}}^{A} \right)\frac{dp}{dt}+\overbrace{\alpha \frac{\partial {{v}_{si}}}{\partial {{x}_{i}}}}^{B}=-\frac{\partial }{\partial {{x}_{i}}}\left( {{v}_{ri}} \right),
\label{eqch14:flmass}
\end{equation}
	
where ${{K}_{g}}$ is the solid grain bulk modulus, and $\alpha$ is the Biot-Willis coefficient ($\alpha =1-K/{{K}_{g}}$ in an ideal, fully interconnected porous media). The bracketed terms $A$ and $B$ represent important couplings from the mechanical system to that of the fluid. All are vital in the HM procedure and without them the equation simplifies to a standard fluid flow equation with fluid compressibility storage in the pressure time derivative. 

\subsubsection*{Solid momentum balance}

We begin with the concept of effective stress,

\begin{equation}
\sigma {{'}_{ij}}={{\sigma }_{ij}}+\alpha p{{\delta }_{ij}},
\end{equation}

for the effective stress, $\sigma '$, and the total stress, $\sigma$; negative in compression. Balance of linear momentum is defined by,

\begin{equation}
\frac{\partial {{\sigma }_{ij}}}{\partial {{x}_{j}}}+{{F}_{i}}=0,
\end{equation}

for the body force, $F=\rho _m \text{g}$ and where $\rho _m =\phi \rho _f + (1-\phi )\rho _s$ is density of the mixture. From the definition of strain, ${{\varepsilon }_{ij}}=\left( \partial {{u}_{i}}/\partial {{x}_{j}}+\partial {{u}_{j}}/\partial {{x}_{i}} \right)/2$, and an arbitrary stress-strain relationship of the form $\mathbf{\sigma }=\mathbf{D\varepsilon }\left( \mathbf{u} \right)$, we write the displacement formulation of mechanical equilibrium (neglecting thermal effects) for isotropic linear elasticity,

\begin{equation}
\frac{\partial }{\partial {{x}_{j}}}\left[ G\frac{\partial {{u}_{i}}}{\partial {{x}_{j}}}+\left( \lambda +G \right)\frac{\partial {{u}_{j}}}{\partial {{x}_{i}}}-\alpha p{{\delta }_{ij}} \right]+{{F}_{i}}=0,
\label{eqch14:mequil}
\end{equation}

where $G$ and $\lambda$ are the Lamè constants. Changes to the fluid system are therefore visited in mechanical equilibrium via the effective stress.

\subsubsection*{Numerical solution scheme}
The numerical solution of Eqs. \ref{eqch14:flmass} and \ref{eqch14:mequil} can be obtained with any convenient method. In these benchmarks, we use a standard Galerkin finite element spatial discretization with time discretization following a generalized first order finite difference scheme, as implemented in OpenGeoSys. Note that Eq. \ref{eqch14:mequil} is an equilibrium equation, and has no time dependency other than that imposed by coupling terms to fluid behavior. The result is a set of coupled linear equations in pressure, $p$, and solid displacement, $u$. The two equations may be solved sequentially and iteratively, or monolithically as a single system. We present results using both solution schemes in the following benchmarks.

%-------------------------------------------------------------------------
%-------------------------------------------------------------------------

\input{chapter_14/subsection_14_1_1}
\input{chapter_14/subsection_14_1_2}
\input{chapter_14/subsection_14_1_3}












