%\section{Continuum mechanics}

The basic idea of continuum mechanics is
that the evolution of a physical system is completely determined by conservation laws,
i.e. basic properties such as mass, momentum, and energy are conserved
during the considered process at all times.
Any physical system can be completely determined by these conservation properties.
In contrast,
other quantities such as pressure or entropy do not obey conservation laws.
The only additional information concerns the consistence of the material
(e.g. fluids, solids, porous medium) in form of constitutive laws.

The concept of conservation means
that the variation of a conservation quantity within a given control volume
is due to the net effect of internal {\it sources} and
of the amount of the quantity which is crossing the boundary surface
of the considered volume - {\it fluxes}.
Sources and fluxes are, in general, dependent of space-time coordinates as well as
on mechanical and thermodynamic factors.
Fluxes result from two contributions:
first due to advective transport by fluid motion and
second due to diffusion/dispersion processes.
Diffusion is always present even when the fluid is at rest.
Diffusion is the tendency towards equilibrium or homogeneity
of a physical system.

%-------------------------------------------------------------------------
%\subsection{Preliminary remarks}
%\label{sec:prelim}

The mechanical description of coupled thermo-hydro-mechanical (THM) processes in porous media is closely associated with the deformation of the solid phase, and the interaction of deformation and flow processes. Each solid material body (including the solid phase of a porous medium) can exhibit different kinds of motion (section \ref{sec:kinematics}):

\begin{itemize}
\item rigid body motion (translation or rotation of the body without changing its volume or shape), and
\item deformation (local relative change of lengths and/or angles referred to neighboring particles, resulting in variations of the shape and/or volume of the material body under consideration).
\end{itemize}

Deformation processes of a porous medium interact with hydraulic processes of the coupled physical system particularly in the following way:

\begin{itemize}
\item effects on the stress state within the solid phase due to pore pressure evolution (with possible risk of rock failure), and
\item variations of the pore size distribution due to the deformation of the solid skeleton, which affect the hydraulic properties, and thus, have an impact on the flow processes in the porous medium.
\end{itemize}

The analysis of deformation processes considered as mechanical response of the material body to the action of applied external forces is one of the objects of mechanics (micromechanics, continuum mechanics, CITATIONS). Porous media distinguish themselves by a sophisticated complex microstructure, whose realistic simulation is extremely challenging, and from a practical point of view generally not efficient. Therefore, continuum mechanics (which is based on the assumption that matter is continuously distributed in space) provides the preferred approaches for the mathematical modeling of deformation processes in porous media. Appropriate models are not based on a physical characterization of the real microstructure, but consider their effects on the physical behavior in a phenomenological manner (section \ref{sec:porous_medium}).

General statements of mechanics, which are independent of the specific material under consideration, refer to the kinematics of motion (shortly described in a following section) and the balance relations (section~\ref{sec:balance_equations}). By contrast, individual material dependent statements refer to the constitutive relations (section.~\ref{sec:m_properties}). Just both the balance relations as well as the constitutive relations constitute a mathematically closed system of equations to solve initial-boundary value problems of mechanics.

\input{211_euler_lagrange}
\input{212_kinematics}
