\chapter{Theory  \today}
\label{cha:theory}
\textit{by Olaf Kolditz, Norbert Böttcher and Uwe-Jens Görke}

Concerning the theoretical background of flow, transport, deformation, and reaction processes in porous media,
there is a considerable amount of monographic literature available \cite{Bea:72,Die:85,Ehl:89,BeaBac:90,Kin:1992,Hel:97,LewSch:98,Boe:00}.
%
The idea of this chapter is to provide a concise brief-as-possible description (compendium-like) of governing equations for thermo-hydro-mechanical / chemical [THM/C] processes in porous media. We will point to literature references rather than giving detailed derivations of the governing equations. This part is the theoretical basis for all benchmarks and examples upcoming in part II and III of this book. We will refer to this part in the examples sections where the working equations are briefly repeated. A list of symbols can be found in the Appendix.

From the mechanical point of view we consider non-isothermal flow of multiple fluid phases (compressible and incompressible fluids) in a deformable thermo-poro-elastic porous medium based on Biot's consolidation concept.
A short introduction to continuum mechanics is given in section \ref{sec:continum_mechanics},
followed by basic conservations principles (section \ref{sec:conservation_principles}) as well as an introduction to theory of porous medium (section \ref{sec:porous_medium}).
%
The followings steps are conducted to derive the general field equations:
\begin{itemize}
 \item Macroscopic balance equations for mass, momentum and energy conservation of porous media (section \ref{sec:balance_equations}),
 \item Constitutive relationships for non-isothermal multiphase flow and deformation processes in porous media, 
 (sections \ref{sec:fluid_properties} and \ref{sec:m_properties}),
 \item Applying the constitutive relationships and introducing physically based simplifications to the balance equations for the derivation of the general field equations. 
 (parts II and III).
\end{itemize}

%-------------------------------------------------------------------------
\section{Continuum mechanics}
\label{sec:continum_mechanics}
\input{21_continuum_mechanics}
%-------------------------------------------------------------------------
\subsection{Conservation principles}
\label{sec:conservation_principles}
\input{22_conservation_principles}
%-------------------------------------------------------------------------
\section{Porous medium}
\label{sec:porous_medium}
\input{23_porous_medium}
%-------------------------------------------------------------------------
\section{Balance equations}
\label{sec:balance_equations}
\input{24_balance_equations}
%-------------------------------------------------------------------------
\section{Fluid properties}
\label{sec:fluid_properties}

The above balance equations derived from first principles (section \ref{sec:balance_equations}) are ''material-less'', i.e. they are valid for any kind of material.
Constitutive relationships are necessary to close the balance equations as well as to specify the properties for fluid flow, heat transfer and mechanical stress/deformation of the specific material under consideration. For determination of material properties laboratory tests have to be conducted. A number of material properties can not be determined directly. This must be done by back analysis using inverse modeling.
%
We organized the description of material properties in two sections: Fluid and mechanical properties (\ref{sec:m_properties}) for THM/C processes in porous media.

\input{25_fluid_properties}
%-------------------------------------------------------------------------
\newpage
\section{Mechanical properties}
\label{sec:m_properties}
\input{26_mechanical_properties}
%-------------------------------------------------------------------------
\input{27_porous_medium_properties}
