The design, implementation and application of a concept for
object-orientated in finite element analysis of multi-field
problems is presented in this paper.
%
The basic idea of this concept is that the underlying governing
equations of porous media mechanics can be classified into
different types of partial differential equations (PDEs). In
principle, equal types of PDEs for diverse physical problems
differ only in material coefficients.
%
Local element matrices and vectors arising from the finite element
discretization of the PDEs are categorized into several types,
regardless of which physical problem they belong to (i.e. fluid flow, mass and
heat transport or deformation processes).
%
Element (ELE) objects are introduced to carry out the local
assembly of the algebraic equations. The object-orientation
includes a strict encapsulation of geometrical (GEO), topological
(MSH), process-related (FEM) data and methods of element objects.
%
Geometric entities of an element such as nodes, edges, faces and
 neighbors are abstracted into corresponding geometric element
objects (ELE-GEO). The relationships among these geometric
entities form the topology of element meshes (ELE-MSH).
%
Finite element objects (ELE-FEM) are presented for the local
element calculations, in which each classification type of the 
matrices and vectors is computed by a unique function. These
element functions are able to deal with different element types
(lines, triangles, quadrilaterals, tetrahedra, prisms, hexhedra) by 
automatically choosing the related element interpolation
functions.
%
For each process of a multi-field problem, only a single instance of 
the finite element object is required. The element objects provide a 
flexible coding environment for multi-field problems with different 
element types. Here, the C++ implementations of the objects are given and 
described in detail.
%
The efficiency of the new element objects is demonstrated by
several test cases dealing with thermo-hydro-mechanical (THM)
coupled problems for geotechnical applications.

\input{31_introduction}
\input{32_sysobj}
%\include{3_symbols}
\input{34_fem}
\input{35_ele_obj}
%\include{example}
%\include{cons}

\begin{itemize}
	\item sparse objects (Wang 2009)
	\item parallelization (Wang et al. 2009)
	\item XFEM (Watanabe 2010)
\end{itemize}