%\chapter{Thermomechanics}
%\textit{by Wenqing Wang, Norihiro Watanabe and Jürgen Hesser}

%makro
%\input{3_symbols}
\newcommand {\StrainT}{ \bm \epsilon}

In this chapter, we consider coupled thermo-mechanical (TM) processes in a porous medium. 
%
%-----------------------------------------------------------
%\section{Theory}
%\subsection*{Heat transport in solids or porous media}
%\label{ssec:heat}
For heat transport problem in any medium, the governing equation is given by
\begin{gather}
 \dens \HC T' = -\nabla \Flux_{\mbox{\tiny T}}+ Q_{\mbox{\tiny T}}(\bm x, t), \,\bm x\in \mathbb R^3
 \label{eq:Tgvn}
\end{gather}
where $\dens$ is medium density, $\HC(T)$ is the specific heat capacity, $ Q_{\mbox{\tiny T}}$ is heat source and  $\Flux_{\mbox{\tiny T}}$ is the heat flux,
which takes
the forms
\begin{equation}
 \Flux_{\mbox{\tiny T}} = -K_e\nabla, T
 \label{eq:tfluxs}
\end{equation}
for solid and
\begin{equation}
 \Flux_{\mbox{\tiny T}} = -K_e\nabla\,T+\poro\sum_{\gamma}^{phase}(\densc\HC^{\gamma}) T \vel, \,\gamma=\mbox{liquid, gaseous }
 \label{eq:tflux1}
\end{equation}
for porous media considering of advective and diffusive fluxes  with $K_e$ the heat conductivity.
For porous  media, the specific heat capacity consists of
 portions of solid, liquid and gaseous phase as
\begin{equation}
  \dens \HC = \sum_{\gamma}^{phase}(\densc\HC^{\gamma})
 \label{eq:tcp}
\end{equation}
where $\gamma$ specifies solid, liquid or  gaseous phase. The  boundary conditions are given by
\begin{equation}
 \Flux_{\mbox{\tiny T}}\cdot\nrl=q^{\mbox{\tiny
 T}}_{\scriptscriptstyle{\Gamma}},\, \mbox{or}\quad
 T=T_{\scriptscriptstyle{\Gamma}}, \,
 \forall\, \Point\, \in \partial \Omega
 \label{eq:tbc1}
\end{equation}
and the initial condition reads
\begin{equation}
 T(\Point, t)=T_0(\Point), \, \forall\, \Point\, \in \Omega
 \label{eq:tini1}
\end{equation}
with $\nrl$, the normal direction at $\Point  \in \partial \Omega$

%\subsection*{Thermal stress}
For mechanical process, we consider the total strain rate $\Delta \StrainT$ can be admissible  decomposed into components such as reversible (elastic),
temperature deduced as
\begin{equation}
\Delta\StrainT=\CT(\Delta\StrainT^{e}-\alpha\,\I \Delta T)
 \label{eq:estrain}
\end{equation}
where $\CT$ is the constitutive tensor, $\alpha$ is the linear thermal expansion coefficient, $\I$ is the identity tensor, and $\Delta T$ is temperature change. With the generalized Hook's law, the total
stress with the thermal effect can be expressed as
\begin{equation}
\Delta\Stress=\CT(\Delta\StrainT-\alpha\,\I \Delta T)
 \label{eq:estress}
\end{equation}
with $\Stress$ is the stress tensor.
%
The volume of a solid is increasing or decreasing with temperature changes. Homogeneous bodies expand evenly in each direction by increasing temperatures. In this case no variation of the stresses occurs. If the deformation of the solid is prevented, the stresses are increasing or decreasing with temperature changes. %(Beitz et al., 1987). 
This phenomenon can be easily calculated by analytical solutions of the Hooke's linear elastic model. The equations of the mechanical behaviour base on the Hooke's law for linear elastic materials:
\begin{eqnarray}
\varepsilon_x & = & \frac{1}{E}\cdot
\left(
\sigma_x\,-\,\nu\cdot
\left(
\sigma_y\,+\,\sigma_z
\right)
\right)\,+\,\alpha\cdot\Delta T
\label{eq61} \\[1.5ex]
\varepsilon_y & = & \frac{1}{E}\cdot
\left(
\sigma_y\,-\,\nu\cdot
\left(
\sigma_x\,+\,\sigma_z
\right)
\right)\,+\,\alpha\cdot\Delta T
\label{eq62} \\[1.5ex]
\varepsilon_z & = & \frac{1}{E}\cdot
\left(
\sigma_z\,-\,\nu\cdot
\left(
\sigma_x\,+\,\sigma_y
\right)
\right)\,+\,\alpha\cdot\Delta T
\label{eq63}
\end{eqnarray}

\noindent
where  $\varepsilon_i$ are strains, $\sigma_i$ are stresses, $E$ is Young's modulus, and $\nu$ is Poisson's ratio. 

%{\small
%with
%\begin{itemize}
%\item[$\varepsilon_i$] -- strains
%\item[$\sigma_i$] -- stresses in Pa
%\item[$E$] -- Young's modulus in Pa
%\item[$\nu$] -- Poisson's ratio
%\item[$\alpha$] -- thermal expansion in K$^{-1}$
%\item[$\Delta T$] -- temperature change in K
%\end{itemize}
%
%indices:
%
%$x,\, y,\, z\;$	- $\;x,\, y,\, z$-direction.
%}


%-----------------------------------------------------------
%\input{PART_III/TM/tm_thermo_elastic_plate_2d}
%\clearpage
%\input{PART_III/TM/tm_thermo_elastic_cube_3d}
%\clearpage
\input{PART_III/TM/tm_homogeneous_material_3d}
%\clearpage
\input{PART_III/TM/tm_composite_material_3d}
%\clearpage
\input{PART_III/TM/tm_hollow_cylinder}
